
% Offer some definitions for technical terms, that lay persons would understand
\usepackage[toc]{glossaries}
\makenoidxglossaries%

\newglossaryentry{videogame}{
    name={videogame},
    description={A game played on a computer system}
}

\newglossaryentry{dlc}{
    name={DLC},
    description={Downloadable Content.
    Some games offer additional game content available to download after the game is published.
    This could be paid for or free content.
    Some may be available at launch time, such as game soundtracks, and other content like new game modes may be made available much later.}
}

\newglossaryentry{}{
    name={},
    description={}
}

\newglossaryentry{asset}{
    name={asset},
    description={A game asset is a digital file which is necessary for the game to function.
    It could be a media file such as a sound effect or graphic.
    It could be a piece of code that implements functionality in the game.}
}
\newglossaryentry{streamedassets}{
    name={streamed assets},
    description={Some games assets are too large to fit on the player's device, so need to be streamed from a server.
    Microsoft Flight simulator has `petabytes'~\cite{microsoft-flight-offline-2021} of data, which is too large to store on one machine.}
}

\newglossaryentry{matchmaking}{
    name={matchmaking},
    description={In a multiplayer game, the game will be responsible for matching players together.
    This could be randomly, based on player ability level, based on their geographic location, or some other factor.
    In a multiplayer game without matchmaking players need to manually connect with another player,
    for example by exchanging connection details over email or social media.}
}

\newglossaryentry{drm}{
    name={DRM},
    description={Digital Rights Management is a form of technology that allows the publisher or manufacturer to impose their own rules or limitations on how a player can interact with their game.
    DRM is often considered a means of imposing restrictions rather than offering rights\cn, and at times has been linked to game breakage\cn.
    It is also present in other digital media such as ebooks\cn or hardware such as printers\cn.}
}

\newglossaryentry{anticheat}{
    name={anti-cheat software},
    description={Anti-cheat software is a particular kind of DRM intended for use in competitive multiplayer games, to ensure that players do not cheat.
    It usually exists as an application separate to the game itself.
    Some games offer players a choice whether to play with it on or off\cite{halo-anticheat-2019}.}
}

\newglossaryentry{unlock}{
    name={unlock},
    description={Some parts of a game might be ``locked'' and unavailable when a player first starts playing a game.
    As the player progresses, parts may be unlocked, for example by spending points earned in the game to unlock a new playable character.
    Some games allow the player to spend real money to unlock things.}
}

\newglossaryentry{save}{
    name={save},
    description={A game save is a description of the state of the game, and the player's progress in games which have a long duration.
    A save may also record things like \glspl{unlock}.
    If the save file is lost, the player will have to start from scratch.
    Saves may be \gls{local} or synchronised \gls{online} between multiple machines.}
}

\newglossaryentry{source}{
    name={source},
    description={Source code is the original text written by a game developer that gets turned into machine code which then runs the game.
    A game can be open or closed source, depending on whether the source code is publicly available or not.}
}

\newglossaryentry{local}{
    name={local},
    description={A local game is one that can be played entirely on one player's machine without an outside internet connection,
    or on multiple player's machines attached to the same network,
    e.g. a WiFi network in one player's house.}
}

\newglossaryentry{online-only}{
    name={online-only},
    description={An online-only game is a game where an internet connection is mandatory.
    If the internet connection is interrupted or poor, the game is unplayable.
    There may be technical reasons why a game has to be online-only,
    or a game may be online-only purely at the choice of the publisher.
    Both single player and multiplayer games can be online-only.}
}

\newglossaryentry{offline}{
    name={offline},
    description={An offline game is a game which requires no networking at all to function.
    Some online games offer configuration options to allow players to play them offline.}
}

\newglossaryentry{online}{
    name={online},
    description={An online game is one which requires internet connectivity to function.
    An online game may need an internet connection for parts of the game.
    If the online connection is missing, those parts of the game may not function correctly.
    Contrast with online-only.}
}

\newglossaryentry{singleplayer}{
    name={singleplayer},
    description={A single player game is intended for one player to play alone.
    Though it may have features such as score leaderboards which allow limited interaction with other players.
    A single game might have both a singleplayer and multiplayer part.}
}

\newglossaryentry{multiplayer}{
    name={multiplayer},
    description={A multiplayer game is intended for multiple players to play together.
    A single game might have both a singleplayer and multiplayer part.
    This could be a \gls{local} or \gls{online} game.
    See also \glc{coop} and \gls{competitive}.}
}

\newglossaryentry{coop}{
    name={cooperative},
    description={A cooperative game is a multiplayer game in which two or more players work together to progress through the game.
    For example, players working together to solve a puzzle.}
}

\newglossaryentry{competitive}{
    name={competitive},
    description={A competitive game is a multiplayer game in which two or more players (alone or in teams) challenge each other, with one party eventually ``winning'' the game.
    For example, a team of players trying to score goals against another team.}
}

\newglossaryentry{client}{
    name={client},
    description={A client is an application that exists on a player's machine which does some functionality on the player machine, but which may connects to other machines on the internet.
    A client may connect to a server.
    Functionality could include drawing graphics onto the player's machine's screen or handling input from a gamepad.}
}

\newglossaryentry{server}{
    name={server},
    description={A server is a machine that runs functionality separate to a client.
    Usually many different clients will connect to it at once.
    It may be run by a player, publisher or a third party.
    It can run functionality, and send the results back to a player's client.
    Functionality could include synchronising the movements of multiple players.}
}

\newglossaryentry{eula}{
    name={EULA},
    description={End User License Agreement. Some games require a user to agree to a license text before they can play the game.
    Accepting or declining the license usually takes place after purchase.
    These documents may or may not be legally binding.}
}

\newglossaryentry{f2p}{
    name={free to play},
    description={Some games may be sold with an up front cost. Others are made available entirely for free.
    Some games are made available for free, but with some content available only to players that purchase it,
    these are called ``free to play''.}
}

\newglossaryentry{update}{
    name={update},
    description={A change made to add, remove or modify some behaviour or content in the game or fix a bug.
    Players may or may not have a choice in whether to accept updates in games.
    Publishers can deliver updates to owners of the game over the internet.}
}

\newglossaryentry{bug}{
    name={bug},
    description={A fault in the design of a videogame or other technology that causes it to not function correctly.
    Could be fixed by an update.}
}
